Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-16 ieee_paper.tex
minor

In this paper, we propose yet a new idea that is applicable to radial
basis functions (RBFs). Their strength is the ability to randomly
distribute points across complex physical domains, and have an
implementation that is independent of dimensionality. RBFs approximate
a function $f(\xvec)\subset \mathbb{R}^d$ sampled at a set of $N$
distinct point locations, $x_j$, by linearly combining translates of a
single radially symmetric function $\phi(r)$, where $r =
\|\xvec-\xvec_{j}\|$ denotes the Euclidean distance (e.g., in 2-D
$\sqrt{(x-x_j)^2+(y-y_j)^2}$) between where the function is evaluated
$\xvec$ and where the RBF is centered $\xvec_{j}$. That is, the
interpolant is $s(\xvec) = \sum_{j=1}^{N} w_j
\phi_i(\|\xvec-\xvec_{j}\|)$. The weights $w_j$ are obtained by
inverting the system
\begin{equation}
\parray{lccr}{
\phi_{11} & \phi_{12} & \cdots & \phi_{1N} \\
\vdots & \ddots & \vdots & \vdots \\
\phi_{n1} & \phi_{n2} & \cdots & \phi_{NN} 
}
\parray{c}{ w_{1} \\ \vdots \\ w_{N} }
=
\parray{c}{ f(\xvec_1) \\ \vdots\\ f(\xvec_N) }. 
\label{eq:rbf}
\end{equation}
where $\phi_{ij} = \phi(\|\xvec_i-\xvec_j\|)$. 
%\end{equation}
%\begin{equation}
%\begin{bmatrix}
%\phi(\|\xvec_{1}-\xvec_{1}\|) & \phi(\|\xvec_{1}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{1}-\xvec_{N}\|) \\
%\vdots & \ddots & \vdots  \\
%\phi (\|\xvec_{n}-\xvec_{1}\|) & \phi(\|\xvec_{n}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{N}-\xvec_{N}\|)
%\end{bmatrix}

The RBF differentiation matrix, $D_N$, is derived by applying the
desired analytic derivative operator $L$ to the RBF interpolant
$s(\xvec)$ above and evaluating it at the point locations. For very
large problems, this is a computationally expensive since the matrix
in (\ref{eq:rbf}) is full and inversion requires O$(N^3)$
operations. To alleviate the cost of this global approach (i.e. using
every node in the domain to calculate the derivative at a given node
$\xvec_i$), RBF-generated finite differences (RBF-FD) have been
derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}. RBF-FD use only
a local set of the $n_z-1$ nearest neighbors to the point $\xvec_i$ to
approximate the derivative. In other words,
$Lf(\xvec_i)=\sum_{j=1}^{n_z}a_jf(\xvec_j)$. The differentiation
weights, $a_j$, are calculated by enforcing that this linear
combination should be exact for RBFs,
$\{\phi(\|\xvec-\xvec_{j}\|)\}_{j=1}^{n_z}$, centered at each of the
node locations $\{\xvec_j\}_{j=1}^{n_z}$ (classical finite differences
(FD) would enforce that it be exact for polynomials instead). Similar
to FD, as the stencil size $n_z$ increases so does the order of the
method.

For a total of $N$ points, there will be $N$ linear systems to solve,
each of size $n_z \times n_z$. Each linear solve produces a row of the
RBF-FD differentiation matrix $D_{n_z}$, resulting in a $N \ times N$
matrix with $n_zN$ nonzero entries. To evaluate the derivative at all
points in the domain, takes the form
$$
  \mathbf{g} = D_{n_z} \mathbf{f}
$$
where $f$ is the source vector of function values and $g$ is the
resulting vector of derivative values. The computation of a single
derivative has been reduced to a SpMV, where each row has $n_z$
nonzeros. Since the sparsity of each row is constant, ELLPACK~\cite{}
becomes the most appropriate matrix compression scheme\cite{}. In
practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
three-dimensional flows. These numbers are similar to what is used in
finite-element codes.

>>>>

In this paper, we propose a new idea that is applicable to radial
basis functions (RBFs). Their strength is the ability to randomly
distribute points across complex physical domains, and have an
implementation that is independent of dimensionality. RBFs approximate
a function $f(\xvec)\subset \mathbb{R}^d$ sampled at a set of $N$
distinct point locations, $x_j$, by linearly combining translates of a
single radially symmetric function $\phi(r)$, where $r =
\|\xvec-\xvec_{j}\|$ denotes the Euclidean distance (e.g., in 2-D
$\sqrt{(x-x_j)^2+(y-y_j)^2}$) between where the function is evaluated
$\xvec$ and where the RBF is centered $\xvec_{j}$. That is, the
interpolant is $s(\xvec) = \sum_{j=1}^{N} w_j
\phi_i(\|\xvec-\xvec_{j}\|)$. The weights $w_j$ are obtained by
inverting the system
\begin{equation}
\parray{lccr}{
\phi_{11} & \phi_{12} & \cdots & \phi_{1N} \\
\vdots & \ddots & \vdots & \vdots \\
\phi_{n1} & \phi_{n2} & \cdots & \phi_{NN} 
}
\parray{c}{ w_{1} \\ \vdots \\ w_{N} }
=
\parray{c}{ f(\xvec_1) \\ \vdots\\ f(\xvec_N) }. 
\label{eq:rbf}
\end{equation}
where $\phi_{ij} = \phi(\|\xvec_i-\xvec_j\|)$. 
%\end{equation}
%\begin{equation}
%\begin{bmatrix}
%\phi(\|\xvec_{1}-\xvec_{1}\|) & \phi(\|\xvec_{1}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{1}-\xvec_{N}\|) \\
%\vdots & \ddots & \vdots  \\
%\phi (\|\xvec_{n}-\xvec_{1}\|) & \phi(\|\xvec_{n}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{N}-\xvec_{N}\|)
%\end{bmatrix}

The RBF differentiation matrix, $D_N$, is derived by applying the
desired analytic derivative operator $L$ to the RBF interpolant
$s(\xvec)$ above and evaluating it at the point locations. For very
large problems, this is a computationally expensive since the matrix
in (\ref{eq:rbf}) is full and inversion requires O$(N^3)$
operations. To alleviate the cost of this global approach (i.e. using
every node in the domain to calculate the derivative at a given node
$\xvec_i$), RBF-generated finite differences (RBF-FD) have been
derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}. RBF-FD uses only
a local set of the $n_z-1$ nearest neighbors to the point $\xvec_i$ to
approximate the derivative. In other words,
$Lf(\xvec_i)=\sum_{j=1}^{n_z}a_jf(\xvec_j)$. The differentiation
weights, $a_j$, are calculated by enforcing that this linear
combination should be exact for RBFs,
$\{\phi(\|\xvec-\xvec_{j}\|)\}_{j=1}^{n_z}$, centered at each of the
node locations $\{\xvec_j\}_{j=1}^{n_z}$ (classical finite differences
(FD) would enforce that it be exact for polynomials instead). Similar
to FD, as the stencil size $n_z$ increases so does the order of the
method.

For a total of $N$ points, there will be $N$ linear systems to solve,
each of size $n_z \times n_z$. Each linear solve produces a row of the
RBF-FD differentiation matrix $D_{n_z}$, resulting in a $N \times N$
matrix with $n_zN$ nonzero entries. To evaluate the derivative at all
points in the domain, takes the form
$$
  \mathbf{g} = D_{n_z} \mathbf{f}
$$
where $f$ is the source vector of function values and $g$ is the
resulting vector of derivative values. The computation of a single
derivative has been reduced to a SpMV, where each row has $n_z$
nonzeros. Since the sparsity of each row is constant, ELLPACK~\cite{}
becomes the most appropriate matrix compression scheme\cite{}. In
practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
three-dimensional flows. These numbers are similar to what is used in
finite-element codes.
